\documentclass[../main.tex]{subfiles}


\begin{document}

\begin{acknowledgments}
First and foremost, I would like to express my sincere gratitude to my advisor, Professor Fabrizio Silvestri. His guidance, insightful feedback, and unwavering support throughout the research process were invaluable. His expertise and encouragement were instrumental in navigating the complexities of this thesis and bringing it to fruition.

I am also deeply indebted to Promo PA Fondazione, particularly its President, Gaetano Scognamiglio, and all its employees. Their willingness to provide opportunities and resources was fundamental to the practical realization of this study. I greatly appreciate their availability and the collaborative spirit they fostered, which significantly enriched my research experience.

This thesis would not have been possible without the essential contribution of the municipalities that collaborated on this project. I extend my heartfelt thanks to the Comune di Lucca and the Comune di Olbia for generously providing the administrative regularity checklists. Access to these real-world documents was crucial for grounding this feasibility study in practical administrative processes and ensuring the relevance of the findings. Their cooperation is a testament to their commitment to innovation in public administration.

I would also like to thank my family and friends for their constant encouragement and patience throughout my studies.

Completing this thesis has been a challenging yet rewarding journey, made possible by the support and contributions of many. Thank you all.

\end{acknowledgments}




\end{document}