\documentclass[../main.tex]{subfiles}
\begin{document}
\section*{Key Findings}
Our experimental results indicate that the automated system achieves an accuracy of approximately 80\% and a balanced accuracy of around 70\% when compared with human evaluations. This demonstrates that, while the LLM-based approach is feasible, human oversight remains essential. Notably, the LLM did not perform as exceptionally as initially hypothesized for a simple question-answering task. We observed that the model's accuracy might be enhanced by incorporating more contextual information about the act, especially in cases involving significant monetary values, where stricter adherence to checklist criteria is imperative. Furthermore, our evaluation revealed that the automated process is considerably faster: while human review of each act takes between 15 to 30 minutes, the LLM completes the evaluation in approximately 2 minutes per act. An important aspect of our findings is that the model tends to err on the side of caution by classifying acts as non-compliant even when they might pass the check. This conservative approach is advantageous, as it ensures that any potential irregularity is flagged for human review rather than mistakenly approving an act.

\section*{Feasibility and Effectiveness}
Our tests have confirmed the feasibility of utilizing LLMs in the domain of administrative regularity control. By conducting a thorough hyperparameter search, particularly focusing on the temperature setting, we found that lower temperature values generally result in more consistent and reliable outputs. Although the current accuracy is promising, it still falls short of replacing human evaluators entirely. Future work will be directed toward enhancing the model's performance, potentially by integrating additional context and refining the prompt engineering strategy. The promising speed and reasonable accuracy suggest that with further improvements, the system could effectively support administrative audits.

\section*{Advantages and Limitations}
The primary advantage of our automated approach is the significant reduction in the time and human resources required for administrative audits. With the LLM handling the initial screening, the process becomes considerably more efficient, which could alleviate the workload on municipal staff. However, several limitations were also identified:
\begin{itemize}
    \item \textbf{Complex Legal Terminology:} The specialized legal language and synonymous terms in administrative acts present challenges for the LLM. There are instances where a human expert might recognize that two terms are synonymous, whereas the model may not.
    \item \textbf{Acronyms and Abbreviations:} Administrative documents often contain acronyms that are not well-defined within the text, which can confuse the model and impact its ability to accurately interpret the act.
\end{itemize}
These limitations underscore the need for further development and integration of domain-specific knowledge into the model.

\section*{Implications for Municipalities}
The integration of an LLM-based auditing system in municipal administrations could have significant positive implications. If further refined, such a system can reduce the administrative burden, speed up the evaluation process, and enhance the overall transparency and accountability of public administration. By ensuring that even minor irregularities are flagged for further review, municipalities can maintain high standards of regulatory compliance. However, it is crucial to retain a human-in-the-loop framework to verify and interpret the model's outputs, especially given the current limitations in handling nuanced legal language.

\section*{Scalability Considerations}
While our methodology has proven effective on a smaller scale, its scalability to other municipalities and contexts requires careful consideration. The approach can be adapted to different types of administrative documents, provided that the documents are pre-processed into a standardized format and the checklists are appropriately customized for local legal and regulatory frameworks. Scaling this system will involve addressing challenges related to the variability in document structure, legal terminology, and the need for more extensive context extraction. Future work should explore these adaptations and consider the integration of additional data sources to further enhance the system's robustness.

\bigskip

\noindent In summary, while the current LLM-based approach demonstrates significant promise in automating administrative regularity controls, it is not yet a complete replacement for human auditors. Nonetheless, it offers a viable and efficient tool that, with further refinement, could substantially improve the auditing processes within public administrations.

\end{document}